---
title: "ECON390 Assignment 1"
author: "Volodymyr Kotov"
output: 
  prettydoc::html_pretty:
    toc: true
    theme: "cayman"
    highlight: "vignette"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

Installing packages:

```{r results='hide', message=FALSE, warning=FALSE}

packages<-c(
  "rlang",
  "dplyr",
  "ggplot2",
  "gridExtra",
  "tidyverse",
  "colourvalues",
  "viridis",
  "usmap",
  "maptools",
  "rgdal",
  "haven",
  "ggrepel"
)

if(!require("pacman")) install.packages("pacman")

pacman::p_load(packages, character.only = T)
sapply(packages, require, character.only = TRUE)

```

Setting up the environment:

```{r results='hide', message=FALSE, warning=FALSE}

setwd("~/Desktop/KSE/Warzone analysis")

```

## 1. Uploading data

### 1a.

```{r results='hide', message=FALSE, warning=FALSE}

data_month_NL <- read.csv("data_month_NL.csv")

```

Creation of light indexes

```{r results='hide', message=FALSE, warning=FALSE}

df1 <- data_month_NL

df1$NLI <- df1$nl_sum_4/(df1$area_sq_km - 0.141*df1$nl_nodata_4)
df1$NLI2 <- df1$nl_mean_4/df1$nl_std_4

```


### 1b.

```{r results='hide', message=FALSE, warning=FALSE}

data_NL_GT <- readRDS("data_NL_GT.rds")

```


### 1c.

```{r results='hide', message=FALSE, warning=FALSE}

Region_by_year_GDP_panel <- read_dta("Region by year GDP panel.dta")

```

## 2. Filtering

```{r results='hide', message=FALSE, warning=FALSE}

#creating a vector of regions of interest
team_6_regs_id <- c(
  "UA_12",
  "UA_53",
  "UA_35",
  "UA_71",
  "UA_05"
  )

#filtering
df2 <- df1 %>% 
  filter(., reg %in% team_6_regs_id)

df3 <- data_NL_GT %>% 
  filter(., reg %in% team_6_regs_id)

df4 <- Region_by_year_GDP_panel %>% 
  filter(., reg %in% team_6_regs_id)

```

## 3. Merging

As we need to select our key variables of interest, we will stick to the following ones:

- data_month_NL:
  - reg, name, year, month, NLI, NLI2, tw_count, area_sq_km
- data_month_NL:
  - reg, year, Labor_index, Holiday_index
- Region_by_year_GDP_panel:
  - reg, year, GDP
  
```{r}

df2 <- df2 %>% 
  select(., c(reg, name, year, month, NLI, NLI2, tw_count, area_sq_km))

df3 <- df3 %>% 
  select(., c(reg, year, Labor_index, Holiday_index))

```


In order to remove a month variable, we will pivot the month variable on the variables of interest:

```{r results='hide', message=FALSE, warning=FALSE}

df2 <- df2 %>%
  pivot_wider(names_from = month, values_from = c(NLI, NLI2, tw_count))

```

Now, we are ready to merge all of the dataframes:

```{r results='hide', message=FALSE, warning=FALSE}

df5 <- list(df2,df3,df4) %>% reduce(left_join, by=c('year','reg'))

```

# 4. Graphs

### Barplots

As we are mostly interested in the impact of war on the economic activity, we build a barplot with the information below:

 - 1st bar: the mean of NLI indexes of all years before 2022
 - 2-3rd bars: NLI in February 2021 and 2022
 - 4-5th bars: NLI in March 2021 and 2022
 - 6-7th bars: NLI in April 2021 and 2022
 - 8-9th bars: NLI in May 2021 and 2022

#### Vinnytska oblast

```{r}

#df5$year<- format(as.Date(df5$year, format="%d/%m/%Y"),"%Y")

df_eda_vin <- df5 %>% 
  filter(., df5$reg == "UA_05" & df5$year > 2011)

Mean_NLI_bw_vin <- c(mean(df_eda_vin$NLI_2[1:(length(df_eda_vin$NLI_2)-1)]),
                    mean(df_eda_vin$NLI_3[1:(length(df_eda_vin$NLI_2)-1)]),
                    mean(df_eda_vin$NLI_4[1:(length(df_eda_vin$NLI_2)-1)]),
                    mean(df_eda_vin$NLI_5[1:(length(df_eda_vin$NLI_2)-1)]))

NLI_2021_vin <- c(df_eda_vin$NLI_2[10],
              df_eda_vin$NLI_3[10],
              df_eda_vin$NLI_4[10],
              df_eda_vin$NLI_5[10])

NLI_2022_vin <- c(df_eda_vin$NLI_2[11],
              df_eda_vin$NLI_3[11],
              df_eda_vin$NLI_4[11],
              df_eda_vin$NLI_5[11])

seq_month <- c("February","March","April","May")

df_vin <- data.frame(Mean_NLI_bw_vin,NLI_2021_vin,NLI_2022_vin,seq_month)

df_vin <- df_vin %>% 
  pivot_longer(
    cols = c("NLI_2021_vin","NLI_2022_vin","Mean_NLI_bw_vin"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_vin$ord <- c(1:12)

df_vin %>% 
  ggplot(., aes(x=reorder(seq_month, ord),y=value, fill = NLI)) +
  geom_bar(stat = "identity",position = "dodge",colour = "black") +
  geom_text(aes(label = round(value,2)), colour = "black", size = 3,
    vjust = 1.5, position = position_dodge(.9)) +
  labs(x="Month",y="NLI",title="Vinnytska oblast NLI", 
       caption="Graph: Barplot")

```

#### Dniprotpetrovska oblast 

```{r}

df_eda_dnp <- df5 %>% 
  filter(., df5$reg == "UA_12" & df5$year > 2011)

Mean_NLI_bw_dnp <- c(mean(df_eda_dnp$NLI_2[1:(length(df_eda_dnp$NLI_2)-1)]),
                    mean(df_eda_dnp$NLI_3[1:(length(df_eda_dnp$NLI_2)-1)]),
                    mean(df_eda_dnp$NLI_4[1:(length(df_eda_dnp$NLI_2)-1)]),
                    mean(df_eda_dnp$NLI_5[1:(length(df_eda_dnp$NLI_2)-1)]))

NLI_2021_dnp <- c(df_eda_dnp$NLI_2[10],
              df_eda_dnp$NLI_3[10],
              df_eda_dnp$NLI_4[10],
              df_eda_dnp$NLI_5[10])

NLI_2022_dnp <- c(df_eda_dnp$NLI_2[11],
              df_eda_dnp$NLI_3[11],
              df_eda_dnp$NLI_4[11],
              df_eda_dnp$NLI_5[11])

df_dnp <- data.frame(Mean_NLI_bw_dnp,NLI_2021_dnp,NLI_2022_dnp,seq_month)

df_dnp <- df_dnp %>% 
  pivot_longer(
    cols = c("NLI_2021_dnp","NLI_2022_dnp","Mean_NLI_bw_dnp"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_dnp$ord <- c(1:12)

df_dnp %>% 
  ggplot(., aes(x=reorder(seq_month, ord),y=value, fill = NLI)) +
  geom_bar(stat = "identity",position = "dodge",colour = "black") +
  geom_text(aes(label = round(value,2)), colour = "black", size = 3,
    vjust = 1.5, position = position_dodge(.9)) +
  labs(x="Month",y="NLI",title="Dnipropetrovska oblast NLI", 
       caption="Graph: Barplot")

```

#### Kirovohradska oblast 

```{r}

df_eda_khd <- df5 %>% 
  filter(., df5$reg == "UA_35" & df5$year > 2011)

Mean_NLI_bw_khd <- c(mean(df_eda_khd$NLI_2[1:(length(df_eda_khd$NLI_2)-1)]),
                    mean(df_eda_khd$NLI_3[1:(length(df_eda_khd$NLI_2)-1)]),
                    mean(df_eda_khd$NLI_4[1:(length(df_eda_khd$NLI_2)-1)]),
                    mean(df_eda_khd$NLI_5[1:(length(df_eda_khd$NLI_2)-1)]))

NLI_2021_khd <- c(df_eda_khd$NLI_2[10],
              df_eda_khd$NLI_3[10],
              df_eda_khd$NLI_4[10],
              df_eda_khd$NLI_5[10])

NLI_2022_khd <- c(df_eda_khd$NLI_2[11],
              df_eda_khd$NLI_3[11],
              df_eda_khd$NLI_4[11],
              df_eda_khd$NLI_5[11])

df_khd <- data.frame(Mean_NLI_bw_khd,NLI_2021_khd,NLI_2022_khd,seq_month)

df_khd <- df_khd %>% 
  pivot_longer(
    cols = c("NLI_2021_khd","NLI_2022_khd","Mean_NLI_bw_khd"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_khd$ord <- c(1:12)

df_khd %>% 
  ggplot(., aes(x=reorder(seq_month, ord),y=value, fill = NLI)) +
  geom_bar(stat = "identity",position = "dodge",colour = "black") +
  geom_text(aes(label = round(value,2)), colour = "black", size = 3,
    vjust = 1.5, position = position_dodge(.9)) +
  labs(x="Month",y="NLI",title="Kirovohradska oblast NLI", 
       caption="Graph: Barplot")

```

#### Poltavska oblast 

```{r}

df_eda_plt <- df5 %>% 
  filter(., df5$reg == "UA_53" & df5$year > 2011)

Mean_NLI_bw_plt <- c(mean(df_eda_plt$NLI_2[1:(length(df_eda_plt$NLI_2)-1)]),
                    mean(df_eda_plt$NLI_3[1:(length(df_eda_plt$NLI_2)-1)]),
                    mean(df_eda_plt$NLI_4[1:(length(df_eda_plt$NLI_2)-1)]),
                    mean(df_eda_plt$NLI_5[1:(length(df_eda_plt$NLI_2)-1)]))

NLI_2021_plt <- c(df_eda_plt$NLI_2[10],
              df_eda_plt$NLI_3[10],
              df_eda_plt$NLI_4[10],
              df_eda_plt$NLI_5[10])

NLI_2022_plt <- c(df_eda_plt$NLI_2[11],
              df_eda_plt$NLI_3[11],
              df_eda_plt$NLI_4[11],
              df_eda_plt$NLI_5[11])

df_plt <- data.frame(Mean_NLI_bw_plt,NLI_2021_plt,NLI_2022_plt,seq_month)

df_plt <- df_plt %>% 
  pivot_longer(
    cols = c("NLI_2021_plt","NLI_2022_plt","Mean_NLI_bw_plt"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_plt$ord <- c(1:12)

df_plt %>% 
  ggplot(., aes(x=reorder(seq_month, ord),y=value, fill = NLI)) +
  geom_bar(stat = "identity",position = "dodge",colour = "black") +
  geom_text(aes(label = round(value,2)), colour = "black", size = 3,
    vjust = 1.5, position = position_dodge(.9)) +
  labs(x="Month",y="NLI",title="Poltavska oblast NLI", 
       caption="Graph: Barplot")

```

#### Cherkaska oblsat

```{r}

df_eda_chr <- df5 %>% 
  filter(., df5$reg == "UA_71" & df5$year > 2011)

Mean_NLI_bw_chr <- c(mean(df_eda_chr$NLI_2[1:(length(df_eda_chr$NLI_2)-1)]),
                    mean(df_eda_chr$NLI_3[1:(length(df_eda_chr$NLI_2)-1)]),
                    mean(df_eda_chr$NLI_4[1:(length(df_eda_chr$NLI_2)-1)]),
                    mean(df_eda_chr$NLI_5[1:(length(df_eda_chr$NLI_2)-1)]))

NLI_2021_chr <- c(df_eda_chr$NLI_2[10],
              df_eda_chr$NLI_3[10],
              df_eda_chr$NLI_4[10],
              df_eda_chr$NLI_5[10])

NLI_2022_chr <- c(df_eda_chr$NLI_2[11],
              df_eda_chr$NLI_3[11],
              df_eda_chr$NLI_4[11],
              df_eda_chr$NLI_5[11])

df_chr <- data.frame(Mean_NLI_bw_chr,NLI_2021_chr,NLI_2022_chr,seq_month)

df_chr <- df_chr %>% 
  pivot_longer(
    cols = c("NLI_2021_chr","NLI_2022_chr","Mean_NLI_bw_chr"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_chr$ord <- c(1:12)

df_chr %>% 
  ggplot(., aes(x=reorder(seq_month, ord),y=value, fill = NLI)) +
  geom_bar(stat = "identity",position = "dodge",colour = "black") +
  geom_text(aes(label = round(value,2)), colour = "black", size = 3,
    vjust = 1.5, position = position_dodge(.9)) +
  labs(x="Month",y="NLI",title="Cherkaska oblast NLI", 
       caption="Graph: Barplot")

```

### Boxplots

We will see our distribution of NLI parameter by months described before to define how unique observations of 2022 are.

#### Vinnytska oblsat

```{r}

df_box_vin <- df5 %>% 
  select(., c("NLI_2","NLI_3","NLI_4","NLI_5")) %>% 
  filter(., df5$reg == "UA_05" & df5$year > 2011 & df5$year < 2022)

df_box_vin <- df_box_vin %>% 
  pivot_longer(
    cols = c("NLI_2","NLI_3","NLI_4","NLI_5"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_box_vin <- df_box_vin[order(df_box_vin$NLI),] 

ggplot(df_box_vin, aes(x=as.factor(NLI), y=value, fill = as.factor(NLI)))+
  geom_boxplot(col='black') + labs(x='Month') +
  scale_fill_viridis_d(option = "D", name="BRC")+
  theme_classic() +
  labs(x="Month",title="Distribution of NLI before war", 
       caption="Graph: Boxplot")

```

#### Dnipropetrovska oblsat

```{r}

df_box_dpr <- df5 %>% 
  select(., c("NLI_2","NLI_3","NLI_4","NLI_5")) %>% 
  filter(., df5$reg == "UA_12" & df5$year > 2011 & df5$year < 2022)

df_box_dpr <- df_box_dpr %>% 
  pivot_longer(
    cols = c("NLI_2","NLI_3","NLI_4","NLI_5"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_box_dpr <- df_box_dpr[order(df_box_dpr$NLI),] 

ggplot(df_box_dpr, aes(x=as.factor(NLI), y=value, fill = as.factor(NLI)))+
  geom_boxplot(col='black') + labs(x='Month') +
  scale_fill_viridis_d(option = "D", name="BRC")+
  theme_classic() +
  labs(x="Month",title="Distribution of NLI before war", 
       caption="Graph: Boxplot")

```

#### Kirovohradska oblsat

```{r}

df_box_khd <- df5 %>% 
  select(., c("NLI_2","NLI_3","NLI_4","NLI_5")) %>% 
  filter(., df5$reg == "UA_35" & df5$year > 2011 & df5$year < 2022)

df_box_khd <- df_box_khd %>% 
  pivot_longer(
    cols = c("NLI_2","NLI_3","NLI_4","NLI_5"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_box_khd <- df_box_khd[order(df_box_khd$NLI),] 

ggplot(df_box_khd, aes(x=as.factor(NLI), y=value, fill = as.factor(NLI)))+
  geom_boxplot(col='black') + labs(x='Month') +
  scale_fill_viridis_d(option = "D", name="BRC")+
  theme_classic() +
  labs(x="Month",title="Distribution of NLI before war", 
       caption="Graph: Boxplot")

```

#### Poltavska oblsat

```{r}

df_box_plt <- df5 %>% 
  select(., c("NLI_2","NLI_3","NLI_4","NLI_5")) %>% 
  filter(., df5$reg == "UA_53" & df5$year > 2011 & df5$year < 2022)

df_box_plt <- df_box_plt %>% 
  pivot_longer(
    cols = c("NLI_2","NLI_3","NLI_4","NLI_5"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_box_plt <- df_box_plt[order(df_box_plt$NLI),] 

ggplot(df_box_plt, aes(x=as.factor(NLI), y=value, fill = as.factor(NLI)))+
  geom_boxplot(col='black') + labs(x='Month') +
  scale_fill_viridis_d(option = "D", name="BRC")+
  theme_classic() +
  labs(x="Month",title="Distribution of NLI before war", 
       caption="Graph: Boxplot")

```

#### Cherkaska oblsat

```{r}

df_box_chr <- df5 %>% 
  select(., c("NLI_2","NLI_3","NLI_4","NLI_5")) %>% 
  filter(., df5$reg == "UA_71" & df5$year > 2011 & df5$year < 2022)

df_box_chr <- df_box_chr %>% 
  pivot_longer(
    cols = c("NLI_2","NLI_3","NLI_4","NLI_5"),
    names_to = "NLI",
    values_to = "value",
    values_drop_na = TRUE
  )

df_box_chr <- df_box_chr[order(df_box_chr$NLI),] 

ggplot(df_box_chr, aes(x=as.factor(NLI), y=value, fill = as.factor(NLI)))+
  geom_boxplot(col='black') + labs(x='Month') +
  scale_fill_viridis_d(option = "D", name="BRC")+
  theme_classic() +
  labs(x="Month",title="Distribution of NLI before war", 
       caption="Graph: Boxplot")

```
As we can see, the real obtained result for this month, which is `r round(df5[13,"NLI_2"],2)` for February falls out of the Q1 to Q3 quantiles. However, the war just began.

### Lineplots

#### Poltavska oblast

```{r message=FALSE, warning=FALSE}

df7 <- df5
df7[is.na(df7)] = 0
df7$mean_NLI <- rowMeans(df7[,5:16])
df7$mean_NLI2 <- rowMeans(df7[,17:28])

ggplot(filter(df7, name == "Poltavska" & year > "2011-02-08"), aes(x = year, y = mean_NLI2)) +
  geom_line(color = "#0099f9", size = 2) +
  geom_point(color = "#0099f9", size = 5) +
  geom_label(
    aes(label = round(mean_NLI2,2)),
     nudge_x = 0.01,
     nudge_y = 0.01
   )

```

#### Vinnytska oblast

```{r message=FALSE, warning=FALSE}

ggplot(filter(df7, name == "Vinnytska" & year > "2011-02-08"), aes(x = year, y = mean_NLI2)) +
  geom_line(color = "#0099f9", size = 2) +
  geom_point(color = "#0099f9", size = 5) +
  geom_label(
    aes(label = round(mean_NLI2,2)),
     nudge_x = 0.01,
     nudge_y = 0.01
   )

```

#### Dnipropetrovska oblast

```{r message=FALSE, warning=FALSE}

ggplot(filter(df7, name == "Dnipropetrovska" & year > "2011-02-08"), aes(x = year, y = mean_NLI2)) +
  geom_line(color = "#0099f9", size = 2) +
  geom_point(color = "#0099f9", size = 5) +
  geom_label(
    aes(label = round(mean_NLI2,2)),
     nudge_x = 0.01,
     nudge_y = 0.01
   )

```

#### Kirovohradska oblast

```{r message=FALSE, warning=FALSE}

ggplot(filter(df7, name == "Kirovohradska" & year > "2011-02-08"), aes(x = year, y = mean_NLI2)) +
  geom_line(color = "#0099f9", size = 2) +
  geom_point(color = "#0099f9", size = 5) +
  geom_label(
    aes(label = round(mean_NLI2,2)),
     nudge_x = 0.01,
     nudge_y = 0.01
   )

```

#### Cherkaska oblast

```{r message=FALSE, warning=FALSE}

ggplot(filter(df7, name == "Cherkaska" & year > "2011-02-08"), aes(x = year, y = mean_NLI2)) +
  geom_line(color = "#0099f9", size = 2) +
  geom_point(color = "#0099f9", size = 5) +
  geom_label(
    aes(label = round(mean_NLI2,2)),
     nudge_x = 0.01,
     nudge_y = 0.01
   )

```


### Piechart

```{r message=FALSE, warning=FALSE}

OtherGDP <- Region_by_year_GDP_panel$GDP[17] - df5$GDP[11] - df5$GDP[24] - df5$GDP[37] - df5$GDP[50] - df5$GDP[63]
Poltavska <- df5$GDP[11]
Vinnytska <- df5$GDP[24]
Dnipropetrovska <- df5$GDP[37]
Kirovohradska <- df5$GDP[50]
Cherkaska <- df5$GDP[63]
name_gdp <- c(OtherGDP, Poltavska, Vinnytska, Dnipropetrovska, Kirovohradska, Cherkaska)

datam <- data.frame(
  Oblast = c("Other", "Poltavska", "Vinnytska", "Dnipropetrovska", "Kirovohradska", "Cherkaska"),
  GDP = name_gdp
  )

pchart <-  ggplot(datam, aes(x="", y = GDP, fill = Oblast)) + 
  geom_bar(stat="identity", width=1, color="black") + 
  coord_polar("y", start=0) +
  theme_void() + 
  theme(legend.position="right")
pchart

```



